---
title: "ISRaD Download"
author: "Grey Monroe"
date: "3/8/2019"
output:
  pdf_document: default
vignette: |
  %\VignetteIndexEntry{Exploring Fractions} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

This brief vingette describes how to download the ISRaD data directly from R using the ISRaD R package.

1. Install (optional)

First as always, you must have installed the ISRaD R package, which provides the function to download the ISRaD data objects. If you already have the ISRaD package you can skip this step.

```{r install, eval=F}
# 1) Install and load the 'devtools' and 'rcrossref' packages
# 'devtools' allows you to install the ISRaD package in its "beta"/development form
# 'rcrossref' is used by our QAQC tool, but sometimes ISRaD won't install without it
install.packages("devtools")
library(devtools) 
install.packages("rcrossref")
library("rcrossref")

# 2) Install package 'ISRaD' from the github repository:
devtools::install_github("International-Soil-Radiocarbon-Database/ISRaD", ref="master")
library(ISRaD) # load the package

```


```{r setup, include=F}
library(ISRaD) # load the package
```

2. Download the data.

The function to download the data is ISRaD.download. This function downloads ISRaD data in .rda format. It also returns this object so you can load the data into R in one step. In the example below, we are downloading the standard ISRaD data object and storing it as a variable called `mydata`.

It accepts two parameters, directory and extra.
`directory` = where you want to downlaod the data object.  
`extra` = whether or not you want to download the ISRaD_extra object.  

```{r download, eval=F}
mydata<-ISRaD.download(directory = "~/Desktop/", extra = F)
```

You can also download the ISRaD_extra dataset which includes additional columns by setting `extra` = T.

```{r extra, eval=F}
mydata_extra<-ISRaD.download(directory = "~/Desktop/", extra = T)
```

