---
title: "ISRaD getdata"
author: "Grey Monroe"
date: "3/8/2019"
output:
  pdf_document: default
vignette: |
  %\VignetteIndexEntry{download} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
---

This brief vingette describes how to load the ISRaD data directly from R using the ISRaD R package.

1. Install (optional)

First as always, you must have installed the ISRaD R package, which provides the function to download the ISRaD data objects. If you already have the ISRaD package you can skip this step.

```{r install, eval=F}
# 1) Install and load the 'devtools' and 'rcrossref' packages
# 'devtools' allows you to install the ISRaD package in its "beta"/development form
# 'rcrossref' is used by our QAQC tool, but sometimes ISRaD won't install without it
install.packages("devtools")
library(devtools) 
install.packages("rcrossref")
library("rcrossref")

# 2) Install package 'ISRaD' from the github repository:
devtools::install_github("International-Soil-Radiocarbon-Database/ISRaD", ref="master")
library(ISRaD) # load the package

```


```{r setup, include=F}
library(ISRaD) # load the package
```

2. Load the data.

The function to load the data is ISRaD.getdata. This function downloads ISRaD data if you dont already have it. It also loads the data into R in one step. It accepts three parameters, directory, dataset, extra.  
`directory` = where the data is located or where you want to download the data.
`dataset` = which dataset you want. This can be the "full" dataset or flattened data objects which are either "flux","interstitial","incubation","fraction",or "layer".
`extra` = whether or not you want to download the ISRaD_extra object.  

In the example below, we are loading the standard ISRaD data object and storing it as a variable called `mydata`.

```{r download, eval=F}
mydata<-ISRaD.getdata(directory = "~/Desktop/", dataset="full", extra = F)
```

