---
title: "ISRaD user guide"
author: "Grey Monroe"
date: "9/11/2018"
output: html_document
---

## How to install the database
Current version is v0.0.1 
for older versions, visit our website.

```{r eval=FALSE}
install.packages("devtools")
install_github("International-Soil-Radiocarbon-Database/ISRaD", ref = "v0.0.1")
library(ISRaD)
```

## How to compile the database
dataset_directory is a directory that contains ISRaD datasets in .xlsx format. Do not include other .xslx files in this directory.
```{r eval=FALSE }
compiled<-compile(dataset_directory = "~/Directory/to/data/", write_report = T, write_out = T, return="flat")
```
What it returns:

| Location      | File name     | Description  |
| ------------- |:---------:| ---------------:|
| dataset_directory/QAQC      | QAQC_\*.txt (\* corresponds to dataset file names) | Report files that identify issues with the dataset files in the dataset_directory |
| col 2 is      | centered      |   $12 |
| zebra stripes | are neat      |    $1 |


## How to subset the database


