---
title: "ISRaD user manual"
author: "Grey Monroe"
date: "9/11/2018"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{ISRaD user manual}
  %\usepackage[UTF-8]{inputenc}
---

## How to install the database
Current version is v0.0.1 
for older versions, visit our website.

```{r eval=FALSE}
install.packages("devtools")
install_github("International-Soil-Radiocarbon-Database/ISRaD", ref = "dev")
library(ISRaD)
```

## How to compile the database
dataset_directory is a directory that contains ISRaD datasets in .xlsx format. Do not include other .xslx files in this directory.
```{r eval=FALSE }
compiled<-compile(dataset_directory = "~/Directory/to/data/", write_report = T, write_out = T, return="flat")
```
What files it creates:

| Description | Location | File name |
| ----------------------------- |:-------:| ------------:|
| Report files that identify issues with the files in the dataset_directory | dataset_directory/QAQC  | QAQC_\*.txt (\* corresponds to dataset file names) |
| Flattened database file    | dataset_directory/database       |   ISRaD_flat.csv |
| List structured database file in the same format as template | dataset_directory/database  |    ISRaD_list.xlsx |
| Log file generated by compile function. Most importatntly, tells you which files passed QAQC. | dataset_directory/database  |   ISRaD_log.txt |
| Summary statistics for datasets compiled into databaser | dataset_directory/database  |   ISRaD_summary.csv |
| QAQC check on compiled database. | dataset_directory/database  |  QAQC_ISRaD_list.txt |



## How to subset the database

## How to visualize the database




